On the eletrodynamic of moving bodies

### 2019-07-03 Article_AS.Rmd
intro version 2

Thousands hectares of Canadian forest are harvested each year and an additional loss is due to insects, diseases and fires [@DRF2000; @NaturalResourcesCanada2019]. It is important to increase tree growth to meet the growing demand of forest industry. The reforestation, when under human control, could ideally be made by using better trees than the former ones, i.e. trees that growth faster, that are more resistant to perturbation and that are of better structural quality. To achieve this task, we must take into account the cambial activity that is governed by environmental conditions and genetics [@Dufour2010; @Rossi2012]. Therefore, genetic selection is one of the available tools and is currently becoming a crucial and increasingly necessary for the development of the canadian forest [@Bonfils2004]. It allows to select trees with characters of interest and to cross them or to reproduce them in order to obtain trees with interesting properties for the development of the forests or the quality of the wood for the human use [@zobel1984; @DRF2000; @nanson2004]. \

Improvement programs by genetic selection are in place for the most important commercial species [@DRF2000; @Mullin2011] including white spruce (Picea glauca (Moench) Voss) which has a significant economic value because of its geographical distribution and abundance [@Eyre1980; @Abrahamson2015]. Spruce wood is valuable as lumber, material for carpentry products and for pulping because of the characteristics of its fibres [@Zhang2008]. In addition, the plasticity of spruce, enable this species to adapt to various types of soil and climatic conditions [@nienstaedt1990]. All these characteristics make white spruce a species of great interest in North America and consequently it is heavily reforested in various silvicultural systems in Canada with more than 150 million of natural an hybrid seedlings planted yearly across almost all forested regions which represents a wide range of soils types and a climatic conditions [@Canadian2009]. \

During the first half of the 20^th^ century, most studies about genetic selection were dedicated on increasing wood volume production [@Dhir1976; @nienstaedt1985; @Li1997; @Koubaa2000] but with the increasing demand of high structural quality products and the fear of losing resource from natural causes, the aspects of wood quality became as important as wood volume [@Franceschini2012]. Therefore, conifer breeding programs are more and more considering wood quality as selection criteria to improve end-use properties. Stem vigor (biologic component) and wood quality (economic component) are two desirable traits that are taken into account in genetic selection. Biologically, the resistance to mortality by vigour is important to ensure to harvest leaving trees. Economically the value of the wood for end-use is important for industries. Several traits were considered to assess wood quality and the density is one of the most widely used indicators and arguably the most important one [@zobel1989; @Saranpaa1994; @Mckendry2002; @Macdonald2002; @Alteyrac2006]. More specifically the homogeneity of density and rings width within the stem has an high importance for stem vigor and wood quality because these two traits are related to it. Homogeneity refers here to a small interannual variation in wood density and rings width.\

In view of the studies cited above, we understand that there might be a strong gain at selecting trees that are prone to produce wood with an homogeneous density to get a resistant and highly valuable forest in the next decades. Consequently, studies have been set up to understand the genetic sources of phenotype variations of wood quality [@Corriveau1971; @Li1993; @Hernandez2001; @Beaulieu2003; @Lenz2013; @McLean2016]. For example, future genetic selections in breeding programs for different species of spruce trees in Quebec will consider wood properties such as wood density [@Beaulieu2009; @Mullin2011]. Hence, it is important to understand where do the variations of density come from in order to better integrated them into genetic selections. In white spruce, there is a typical low frequency radial pattern of variation from pith to bark [@Moore2011] with a density that decreases during the first years of growth and then an increasing from the beginning of the formation of mature wood [@panshin1980]. These variations are under genetic control and are consequently moderately to highly heritable [@Corriveau1991; @Yanchuk1993; @Ivkovich2002a; @Ivkovich2002b; @Beaulieu2006]. This makes this trait attractive for genetic improvement. In addition to these low-frequency variations, there are high frequency inter-annual variations [@Nicault2010] that are influenced mainly by events to the tree such as the soil water availability and climatic conditions [@Nicault2010] that happen at high frequency. These variations are not under genetic control but under climate control meaning that climatic conditions also affect wood density. Some climatic factors such as drought can cause stress in trees and cause a variation in wood properties, including density, ring width and conductivity. Long-term growth analyses have shown that density may even increases with the temperature, the atmospheric CO~2~ content and the nitrogen deposition [@becker1995; @huang2007; @zaehle2014]. Wood density can also be negatively correlated with growth. [@zhang1995]. Therefore, with the aim of producing wood of homogeneous density it is essential to take climate factors into account in addition to genetic.\

 In fact, the density is correlated with the growth of the tree [@corriveau1991] and it has been observed that trees with a small variation in growth and therefore in density are also the most resistant to mortality [@ogle2000; @suarez2004; @hartmann2008]. For example, @suarez2004 found that at the tree scale, individuals with variable growth were more prone to die of drought than more regularly growing trees. These studies have shown that the steady growth and therefore the small variation in the density of the wood allows the tree to be more vigorous. With regard to the economical context, variation of density within a piece of wood may lead to deformation during transformation and inhomogeneous machining properties. Indeed, wood density homogeneity affects the performance of saw timber, the pulp yield and calorific values for biomass applications for bio energy [@Hernandez2001; @Macdonald2002; @Mckendry2002]. For example, the low variation in wood density is appreciated for sawnwood because sawyers place a high value on the homogeneity of wood density [Hassegawa et al (pending approval); @X]. It is therefore important from both a biological and an end-use points of view to have a wood with a uniform density within the stem. \

A significant variation of the climate can cause a variation of the properties of the wood over several years. For example, several studies have shown that after the great drought of 2003 in Europe, the growth of forests has been affected over several years [@ciais2005; @granier2007; @reichstein2007]. The ability of trees to tolerate these changes in a tree species or within a species is often related to the interaction between genotype and environment. For example, studies have suggested a possible link between low sensitivity to drought and genotype [@ferrell1966; @teskey1987]. This link is poorly controlled and therefore poorly known. It is in context that this study fits to try to advance the knowledge in this direction.\

In this study we quantified and compared heterogeneity of wood density from trees of known parents in order to figure out if some individuals produce more heterogeneous wood than others. Knowing the parents of each tree we tried to characterize the genetic contribution to the heterogeneity by comparing two growth sites we tried to characterize the environmental contribution to homogeneity. We ultimately aim to select and reproduce white spruce trees families of interest that are both low-sensitivity to climate change and genetically prone to make homogeneous wood. Thus, the most homogeneous families will be identified for reforestation and identify variables to consider in breeding of growing trees with less variable wood.

>>>>

Thousands hectares of Canadian forest are harvested each year and an additional loss is due to insects, diseases and fires [@DRF2000; @NaturalResourcesCanada2019]. It is important to increase tree growth to meet the growing demand of forest industry. The reforestation, when under human control, could ideally be made by using better trees than the former ones, i.e. trees that growth faster, that are more resistant to perturbation and that are of better structural quality. To achieve this task, we must take into account the cambial activity that is governed by environmental conditions and genetics [@Dufour2010; @Rossi2012]. Therefore, genetic selection is one of the available tools and is currently becoming a crucial and increasingly necessary for the development of the canadian forest [@Bonfils2004]. In forestry, genetic selection is a process allows to select trees with characters of interest and to cross them or to reproduce them in order to obtain trees with interesting properties for the development of the forests or the quality of the wood for human use [@zobel1984; @DRF2000; @nanson2004]. \

During the first half of the 20^th^ century, most studies about genetic selection were dedicated on increasing wood volume production [@Dhir1976; @nienstaedt1985; @Li1997; @Koubaa2000] but with the increasing demand of high structural quality products and the fear of losing resource from natural causes, the aspects of wood quality became as important as wood volume [@Franceschini2012]. Hence, conifer breeding programs are more and more considering wood quality as selection criteria to improve end-use properties. For example, improvement programs by genetic selection are in place for white spruce *(Picea glauca [Moench] Voss)* [@DRF2000; @Mullin2011] which has a significant economic value because of its geographical distribution and abundance [@Eyre1980; @Abrahamson2015] in North America. Spruce wood is valuable as lumber, material for carpentry products and for pulping because of the characteristics of its fibres [@Zhang2008]. In addition, the plasticity of spruce, enable this species to adapt to various types of soil and climatic conditions [@nienstaedt1990]. All these characteristics make white spruce a species of great interest in North America and consequently it is heavily reforested in various silvicultural systems in Canada with more than 150 million of natural an hybrid seedlings planted yearly across almost all forested regions which represents a wide range of soils types and a climatic conditions [@Canadian2009]. In Quebec, several generations of seed orchards have already been established in white spruce. Thus, the current objectives of genetic selection for this species are to select trees with a higher wood quality  because white spruce has already reached a level considered interesting for traits such as rapid growth and volume of wood. \

@mitchell1961 defined the quality of the wood as follows: "Wood quality is the resultant of physical and chemical characteristics possessed by a tree or a part of a tree that enable it to meet the property requirements for different end products". it is composed mainly of two components: economical and biological (stem vigor). These two components are threfore two desirable traits that are taken into account in genetic selection. Biologically, resistance to mortality by vigor is important to ensure the survival of the tree. Economically, the value of the wood for end-use is important for industries. Several traits were considered to assess wood quality and the density is one of the most widely used indicators and arguably the most important one [@zobel1989; @Saranpaa1994; @Mckendry2002; @Macdonald2002; @Alteyrac2006]. More specifically the low interannual variability of wood density and rings width has an high importance for stem vigor and wood quality because these two traits are related to it. In fact, the density is correlated with the growth of the tree [@corriveau1991] and it has been observed that trees with a low variation in growth and therefore in density are also the most resistant to mortality [@ogle2000; @suarez2004; @hartmann2008]. For example, @suarez2004 found that at the tree scale, individuals with variable growth were more prone to die of drought than more regularly growing trees. These studies have shown that the steady growth and therefore the low interannual variation wood density allows the tree to be more vigorous and therefore less prone to related mortality and environmental conditions. With regard to the economical context, interannual variation of density within a piece of wood may lead to deformation during transformation and inhomogeneous machining properties. Indeed, low interannual variability of wood density affects the performance of saw timber, the pulp yield and calorific values for biomass applications for bioenergy [@Hernandez2001; @Macdonald2002; @Mckendry2002]. For example, the low interannual variation in wood density is appreciated for sawnwood because sawyers place a high value on the low interannual variation wood density [Hassegawa et al (pending approval); @X]. It is therefore important from both a biological and an end-use points of view to have a wood with a low interannual density variation within the stem. We understand that there might be a strong gain at selecting trees that are prone to produce wood with a low interannual variability of wood density to get a resistant and highly valuable forest in the next decades. For example, future genetic selections in breeding programs for different species of spruce trees in Quebec will consider wood properties such as wood density [@Beaulieu2009; @Mullin2011]. For wood quality, one of the characteristics that one wishes to select is the amplitude of the interannual variation of wood density, provided that it has sufficiently interesting heritability for a genetic selection. \

Heritability is the probability that a trait of an individual is hereditarily transmitted by genetic factors. It is a statistical datum that serves to evaluate the part of genetic factors and that of environmental factors in the construction of the phenotype of a given population. Several studies have been set up to understand the genetic sources of phenotypic variations of wood quality traits and their heritability [@Corriveau1971; @Li1993; @Hernandez2001; @Beaulieu2003; @Lenz2013; @McLean2016]. Thus, previous studies have shown that wood density is a trait under strong genetic control therefore strong heritability [@Corriveau1991; @Yanchuk1993; @Ivkovich2002a; @Ivkovich2002b; @Beaulieu2006]. It has been proven that the wood density is under genetic control, but we can ask ourselves the question of its amplitude of interannual variation. The interannual variation of wood density comes from two sources: the low-frequency radial profile and the high-frequency radial profile [@Nicault2010]. The low-frequency radial profile is the radial variation of the density of pith-to-bark with a  high density during the first years of tree growth  decreases to a certain level and begins to increase again. In white spruce, this radial profile is the radial type II profile described by [@panshin1980]. These variations are under genetic control and are consequently moderately to highly heritable [@Corriveau1991; @Yanchuk1993; @Ivkovich2002a; @Ivkovich2002b; @Beaulieu2006] because this functioning is internal to the tree and is transmitted to future generations. The high-frequency radial profile is the amplitude of the interannual variation of wood density. We did not find any studies on heritability of amplitude of interannual variation of wood density in white spruce.\

Studies have linked annual environmental conditions such as soil water availability, climatic conditions that occur at a high frequency, and interannual variability wood density [@becker1995; @huang2007; @Nicault2010; @zaehle2014]. Indeed, long-term growth analyzes have shown a link between wood density and local annual conditions such as summer temperature in white spruce [@darrigo1992] and other species [@becker1995; @huang2007; @zaehle2014]. However, it seems to have some genetic control in the density sensitivity to environmental conditions. For example, studies have suggested a possible link between low sensitivity to drought and genotype [@ferrell1966; @teskey1987]. These studies show that although the interannual variation in density is under environmental control, its amplitude may be controlled by a genotype-environment interaction. Therefore, a question arises as to whether the amplitude of interannual variation is primarily under genetic control (heritability). Thus, in the context of climate change, it is crucial to know the level of heritability of the amplitude of interannual variation in wood density for the genetic selection of individuals with low amplitude of density variation. \

In this study, we evaluated the amplitude of the interannual variation of wood density in white spruce and its share of heritability in the interannual conditions of the environment. We therefore compare two growth sites containing genotyped trees in order to characterize the genetic (heritability) and environmental contribution to the interannual variation of wood density. Thus, we hope to select and reproduce white spruce trees families of interest that are both low-sensitivity to climates changes and genetically prone to make low amplitude of interannual variation of wood density. If the amplitude of variation in wood density is under genetic control, tree families with the smallest variation amplitudes will be identified for reforestation.
